Sequential Composition for Relaxed Memory: Pomsets with Predicate Transformers

### 2021-02-11 paper/arm.tex
Draft of Fork-Join
Sorted out bib

\begin{example}
  \label{ex:dgr1}
  The following execution is allowed by \armeight{}, but disallowed by
  \refdef{def:pomsets-ra}.  The coherence order between the writes can be
  witnessed by a separate thread, which we have elided.
  \begin{gather*}
    \PW{x}{2}\SEMI 
    \PR[\mRA]{x}{r}\SEMI
    \PW{y}{1} \PAR
    \PW{y}{2}\SEMI
    \PW[\mRA]{x}{1}
    \\
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DW{x}{2}}{}
        \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
        \event{c}{\DW{y}{1}}{right=of b}
        \event{d}{\DW{y}{2}}{right=2.5em of c}
        \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
        \rf{a}{b}
        \sync{b}{c}
        \wk{c}{d}
        \sync{d}{e}
        \wk[out=-165,in=-15]{e}{a}
      \end{tikzinline}}
  \end{gather*}
  Under \EGC{}, this is explained by dropping the order
  $\DRP[\mRA]{x}{2}\xsync\DWP{y}{1}$, because $\DRP[\mRA]{x}{2}$ is fulfilled
  by a relaxed write in the same thread.
  \begin{gather*}
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DW{x}{2}}{}
        \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
        \event{c}{\DW{y}{1}}{right=of b}
        \event{d}{\DW{y}{2}}{right=2.5em of c}
        \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
        \rf{a}{b}
        % \sync{b}{c}
        \wk{c}{d}
        \sync{d}{e}
        \wk[out=-165,in=-15]{e}{a}
      \end{tikzinline}}  
  \end{gather*}

@@

\begin{definition}[\xCO/\xRASC/\xDGR/\xRRD]
  \label{def:pomsets-arm}
  Update \refdef{def:pomsets-trans} to: % (\ref{L4} unchanged):
  \begin{enumerate}
  \item[\ref{S3})]
    $\labelingForm(\aEv)$ implies $\QS{\aLoc}{\amode}\land\aExp{=}\aVal$,
  \item[\ref{L3})]
    $\labelingForm(\aEv)$ implies $\QL{\aLoc}{\amode}$.
  \item[\ref{T3})]
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\TRUE/\RW]$ if $\labelingAct_1(\aEv)$ is a write,
    \\
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\FALSE/\RW]$ otherwise.
  \end{enumerate}
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies
    \begin{math}
      %(\Qw{\aLoc}\limplies\aExp{\neq}\aVal)\land
      \bForm %\noSUB{[\aExp/\aLoc]},
      [(\Qw{\aLoc}\land\aExp{=}\aVal)/\Qw{\aLoc}]
      \DS{\aLoc}{\amode}
    \end{math}
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies
    \begin{math}
      %\lnot\Qw{\aLoc}\land 
      \bForm %\noSUB{[\aExp/\aLoc]}.
      [\FALSE/\QS{\aLoc}{\amode}]
      \DS{\aLoc}{\amode}
    \end{math}
  \item[\ref{L4})]
    $\aTr{\bEvs}{\bForm}$ implies $\aVal{=}\aReg\limplies\bForm$, 
  \item[\ref{L5})]
    % $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land(\RW\limplies(\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm)$,
    % \item[\ref{L6})]
    $\aTr{\cEvs}{\bForm}$ implies
    \begin{math}
      % \lnot\Qx{\aLoc} \land
      \DL{\aLoc}{\amode}{} \land (\RW\limplies\bForm[\FALSE/\QL{\aLoc}{\amode}]).
    \end{math}
  \end{enumerate}  
\end{definition}

>>>>

\begin{example}
  \label{ex:dgr1}
  The following example is from \citeauthor{anton} \cite{anton}.  It is
  allowed by \armeight{}, but disallowed by \refdef{def:pomsets-ra}.  The
  coherence order between the writes can be witnessed by a separate thread,
  which we have elided.
  \begin{gather*}
    \PW{x}{2}\SEMI 
    \PR[\mRA]{x}{r}\SEMI
    \PW{y}{1} \PAR
    \PW{y}{2}\SEMI
    \PW[\mRA]{x}{1}
    \\
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DW{x}{2}}{}
        \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
        \event{c}{\DW{y}{1}}{right=of b}
        \event{d}{\DW{y}{2}}{right=2.5em of c}
        \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
        \rf{a}{b}
        \sync{b}{c}
        \wk{c}{d}
        \sync{d}{e}
        \wk[out=-165,in=-15]{e}{a}
      \end{tikzinline}}
  \end{gather*}
  Under \EGC{}, this is explained by dropping the order
  $\DRP[\mRA]{x}{2}\xsync\DWP{y}{1}$, because $\DRP[\mRA]{x}{2}$ is fulfilled
  by a relaxed write in the same thread.
  \begin{gather*}
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DW{x}{2}}{}
        \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
        \event{c}{\DW{y}{1}}{right=of b}
        \event{d}{\DW{y}{2}}{right=2.5em of c}
        \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
        \rf{a}{b}
        % \sync{b}{c}
        \wk{c}{d}
        \sync{d}{e}
        \wk[out=-165,in=-15]{e}{a}
      \end{tikzinline}}  
  \end{gather*}

@@

\begin{definition}[\xCO/\xRASC/\xDGR/\xRRD]
  \label{def:pomsets-arm}
  Let $[\aForm/\Q{}]$ be the substitution $[\aForm/\Qr{*}][\aForm/\Qw{*}][\aForm/\Qsc]$.  
  Update \refdef{def:pomsets-trans} to: % (\ref{L4} unchanged):
  \begin{enumerate}
  \item[\ref{S3})]
    $\labelingForm(\aEv)$ implies $\QS{\aLoc}{\amode}\land\aExp{=}\aVal$,
  \item[\ref{L3})]
    $\labelingForm(\aEv)$ implies $\QL{\aLoc}{\amode}$.
  \item[\ref{T3})]
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\TRUE/\RW]$ if $\labelingAct_1(\aEv)$ is a write,
    \\
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\FALSE/\RW]$ otherwise.
  \end{enumerate}
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies
    \begin{math}
      %(\Qw{\aLoc}\limplies\aExp{\neq}\aVal)\land
      \bForm %\noSUB{[\aExp/\aLoc]},
      [(\Qw{\aLoc}\land\aExp{=}\aVal)/\Qw{\aLoc}]
      \DS{\aLoc}{\amode}
    \end{math}
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies
    \begin{math}
      %\lnot\Qw{\aLoc}\land 
      \bForm %\noSUB{[\aExp/\aLoc]}.
      [\FALSE/\QS{\aLoc}{\amode}]
      \DS{\aLoc}{\amode}
    \end{math}
  \item[\ref{L4})]
    $\aTr{\bEvs}{\bForm}$ implies $\aVal{=}\aReg\limplies\bForm$, 
  \item[\ref{L5})]
    % $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land(\RW\limplies(\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm)$,
    % \item[\ref{L6})]
    $\aTr{\cEvs}{\bForm}$ implies
    \begin{math}
      % \lnot\Qx{\aLoc} \land
      \DL{\aLoc}{\amode}{} \land (\RW\limplies\bForm[\FALSE/\QL{\aLoc}{\amode}]).
    \end{math}
  \end{enumerate}  
\end{definition}

### 2021-02-11 paper/intro.tex
Sorted out bib

Because it is closely related, we expect that the memory-model results of
\cite{DBLP:journals/pacmpl/JagadeesanJR20} apply to our model, including
compositional reasoning for temporal safety properties and {local} \drfsc.
In \textsection\ref{sec:arm}, we provide an alternative proof strategy for
efficient compilation to \armeight{}, which improves upon that of
\cite{DBLP:journals/pacmpl/JagadeesanJR20} by using a recent alternative
characterization of \armeight{}.

>>>>

Because it is closely related, we expect that the memory-model results of
\cite{DBLP:journals/pacmpl/JagadeesanJR20} apply to our model, including
compositional reasoning for temporal safety properties and {local} \drfsc{}
as in \cite{Dolan:2018:BDR:3192366.3192421,DBLP:conf/ppopp/DongolJR19}.
In \textsection\ref{sec:arm}, we provide an alternative proof strategy for
efficient compilation to \armeight{}, which improves upon that of
\cite{DBLP:journals/pacmpl/JagadeesanJR20} by using a recent alternative
characterization of \armeight{}.
